\documentclass[12pt,a4paper,oneside]{abntex2}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{esint}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfig}
\graphicspath{{fig/}}
\providecommand{\norm}[1]{\lVert#1\rVert}
\usepackage{empheq}

%\usepackage[alf ,abnt-etal-cite=2 , abnt-year-extra-label=yes , abnt-etal-list=0, abnt-etal-text=it]{abntex2cite}

%capa
\autor{DAVID DA COSTA DE PINHO}% introduz nome do autor
\titulo{FUNDAMENTOS DE MAGNETO-ELASTICIDADE}
\data{2018}
\local{MACA\'E}

%anverso da folha de rosto
\preambulo{Monografia apresentada ao Centro de Ci\^encia e Tecnologia da Universidade Estadual do Norte Fluminense, como parte das exig\^encias para aprova\c{c}\~ao no Exame de Qualifica\c{c}\~ao.}
\orientador{Viatcheslav Ivanovich Priimenko, Ph.D}
\tipotrabalho{monografia}

\begin{document}

\imprimircapa

\imprimirfolhaderosto%anverso

%{\ABNTEXchapterfont\Large\textsc{\imprimirautor}} - para aumentar a fonte e colocar em caixa alta

%importante dar enter para pular linha entre os tipos de textos

\begin{folhadeaprovacao}

\begin{center}
{\ABNTEXchapterfont\Large\bfseries\imprimirtitulo}\\\vspace{1cm}
{\ABNTEXchapterfont\Large\textsc{\imprimirautor}}

\end{center}

\vspace{1cm}
\hspace{.45\textwidth} \begin{minipage}{.5\textwidth}
\imprimirpreambulo
\end{minipage}

\vspace{1cm}
Trabalho aprovado em $\,7\,$ de Agosto de \imprimirdata.\\

\vspace{1cm}
Comiss\~ao Examinadora:\\
\assinatura{Prof. Marcia Miranda Azeredo, Ds.c - UENF}
\assinatura{Prof. André Duarte Bueno, Ds.c - UENF}
\assinatura{Prof. Fernando Diogo de Siqueira, Ds.c - UENF}
\assinatura{Prof. \imprimirorientador - UENF}

\begin{center}
\vfill
{\large\imprimirlocal}
\par
{\large\imprimirdata}
\end{center}

\end{folhadeaprovacao}

\begin{center}
{\ABNTEXchapterfont\Large\imprimirtitulo}\\\vspace{1cm}

\end{center}

\begin{resumo}
O objetivo principal desta obra \'e apresentar os fundamentos das teorias f\'isicas e matem\'atica relacionados ao efeito magneto-el\'astico, bem como um recondicionamento do modelo usado para descrever tal efeito. Inicialmente, apresentamos uma revis\~ao dos principais fen\^omenos pesquisados atualmente, envolvendo propaga\c{c}\~ao de ondas el\'asticas e eletromagn\'eticas. A mec\^anica do cont\'inuo e o eletromagnetismo s\~ao as duas teorias f\'isicas que descrevem a magneto-elasticidade, e introduzimos os principais conceitos dessas \'areas necess\'arios aos propositos desta monografia. Estudamos como ocorre o acoplamento entre ondas el\'asticas e eletromagn\'eticas bem como sua propaga\c{c}\~ao na subsuperf\'icie terrestre, atrav\'es de camadas estratigr\'aficas. Finalizamos com uma reescrita do modelo desse acoplamento de forma que possamos desenvolv\^e-lo analiticamente e produzir uma solu\c{c}\~ao num\'erica em trabalhos posteriores.

\vspace{\onelineskip} 
\noindent 
\textbf{Palavras-chave}: Magneto-Elasticidade, Mec\^anica do Cont\'inuo, Equa\c{c}\~oes de Maxwell.
\end{resumo}

\begin{center}
{\ABNTEXchapterfont\Large FOUNDATIONS OF MAGNETO-ELASTICITY}\\\vspace{1cm}
\end{center}

\begin{resumo}[Abstract]
\begin{otherlanguage*}{english}
The main purpose of this paper is present the foundations of physical and mathematical theories about magneto-elastic effect, and a reconditioning model to describe such effect. Firstly, we present a review of some phenomena found in current surveys, involving the propagations of elastic and eletromagnetics waves. Continuous mechanics and eletromagnetism describe magneto-elasticity, and we introduce a lots of required concepts from these areas to achieve the monography's goals. We also study how occurs the coupled propagation of elastic and eletromagnetics waves in Earth's subsurface, through stratigraphics layers. We finished rewriting the coupling's model in such manner that allow us to develope it analiticaly and to build a numerical solver in further works.

\vspace{\onelineskip} 
\noindent 
\textbf{Keywords}: Magneto-Elasticity, Continuous Mechanics, Maxwell's Equations.

\end{otherlanguage*}
\end{resumo}

\listoffigures

\newpage

\tableofcontents

\textual

\input{introducao_eletro_elasti}
\input{revisao_bibliografica_1}
\input{fund_eletro}
\input{fund_elasti}
\input{acopla_magneto_elastico}
\input{recondi-acoplamento}
\input{conclusoes}

\bibliographystyle{plainnat}
%\bibliography{referencias}

\begin{thebibliography}{25}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi


\bibitem[Cukavac(2008)]{Cukavac_2008}
M.~S. Cukavac.
\newblock Seismomagnetic insvestigations in kopaonik area.
\newblock \emph{MGB}, 2008.

\bibitem[Eringen(1963)]{eringen_1963}
J.W.~Dunkin e~A.C.~Eringen.
\newblock On the propagation of waves in an electromagnetic elastic solid.
\newblock \emph{International Journal of Engineering Science}, 1, 1963.


\bibitem[Tromp(1998)]{dahlem}
F.~A.~Dahlem e~J.~Tromp.
\newblock \emph{Theoretical Global Seismology}.
\newblock Princeton University Press, 1998.

\bibitem[Soboleva(1997)]{Mikhailenko_1997}
B.~G.~Mikhailenko e~O.~N.~Soboleva.
\newblock Mathematical modeling of seismomagnetic efects arising in the seismic
  wave motion in the earth's constant magnetic field.
\newblock \emph{Applied Mathematics Lettures}, 10\penalty0 (3):\penalty0
  47--51, 1997.

\bibitem[Pilipenko(1997)]{surkov_97}
V.~V.~Surkov e~V.~A.~Pilipenko.
\newblock Magnetic effects due to earthquakes and underground explosions: a
  review.
\newblock \emph{Annali di Geofisica}, XL\penalty0 (2):\penalty0 227--239, 1997.

\bibitem[Eringen(1962)]{Eringen_1962}
A.~C. Eringen.
\newblock \emph{Nonlinear Theory of Continuous Media}.
\newblock McGraw-Hill New York, 1962.

\bibitem[Griffiths(1999)]{griffiths}
D.~J. Griffiths.
\newblock \emph{Introduction to Electrodynamics}.
\newblock Prentice-Hall, 1999.

\bibitem[Guglielmi(1986a)]{guglielmi_86a}
A.~V. Guglielmi.
\newblock Magnetoelastic waves.
\newblock \emph{Izv. Akad. Nauk SSSR, Fizika Zemli}, 7\penalty0 (112), 1986a.

\bibitem[Guglielmi(1986b)]{guglielmi_86b}
A.~V. Guglielmi.
\newblock Excitation of oscillations of the electromagnetic field by elastic
  waves in the conducting body.
\newblock \emph{Geomagn. Aeron.}, 27\penalty0 (3):\penalty0 467--470, 1986b.

\bibitem[Jackson(1999)]{jackson_classical_1999}
J.~D. Jackson.
\newblock \emph{Classical electrodynamics}.
\newblock Wiley, New York, {NY}, 3rd ed., 1999.


\bibitem[Knopoff(1955)]{Knopoff_1955}
E.~L. Knopoff.
\newblock The interaction between elastic waves motions and magnetic field in
  electrical conductors.
\newblock \emph{J. Geophys. Res.}, 60\penalty0 (4):\penalty0 617--629, 1955.

\bibitem[Yerzhanov(1985)]{yerzhanov_85}
T.~E. Nasynbaev e A. V.~Bushuev L.~S.~Yerzhanov, A. K.~Kurskeev.
\newblock Geomagnetics observations during the massa experiments.
\newblock \emph{Izv. Akad. Nauk SSSR, Fizika Zemli}, 11:\penalty0 80--82, 1985.

\bibitem[Liu(2002)]{liu}
I.~S. Liu.
\newblock \emph{Continuum Mechanics}.
\newblock Spring-Verlag, Berlim-Heidelberg, 2002.

\bibitem[Sadovsky(1980)]{Sadovsky_1980}
M.A. Sadovsky.
\newblock Electro magnetic precursors of earthquakes.
\newblock \emph{Dokl. Acad. Nauka}, 1980.

\bibitem[Slawinski(2007)]{slawinski}
M.~A. Slawinski.
\newblock \emph{Waves And rays in elastic continua}.
\newblock World Scientific Publishing Company, 2 ed, 2007.


\bibitem[Sommerfeld(1952)]{sommerfeld_52}
A.~Sommerfeld.
\newblock Electrodynamics.
\newblock \emph{Academic Press}, 1952.

\bibitem[Stacey(1964)]{stacey_64}
F.~D. Stacey.
\newblock Seismo-magnetic effect.
\newblock \emph{Pure Applied Geophysics}, 58\penalty0 (11):\penalty0 5--23,
  1964.

\bibitem[Surkov(1989a)]{surkov_89a}
V.~V. Surkov.
\newblock Local changes in geomagnetics and geoeletrics fields under rocks
  deformation near the earth surface.
\newblock \emph{Izv. Akad. Nauk SSSR, Fizika Zemli}, 5:\penalty0 91--96, 1989a.

\bibitem[Surkov(1989b)]{surkov_89b}
V.~V. Surkov.
\newblock Distortion of external magnetic field by a longitudinal acoustic
  wave.
\newblock \emph{Magnetic Hydro-Dynamics}, 2:\penalty0 9--12, 1989b.

\bibitem[Anisimov(1985)]{Anisimov_1985}
E.A. Ivanov M.V. Pedanov N.N.Rusakov V.A. Troizhkya e V.E.~Goncharov
  S.V.~Anisimov, M.B.~Gokhberg.
\newblock Short period oscillations of electromagnetic field of the earth after
  explosion.
\newblock \emph{Dokl. Acad. Nauka}, 281\penalty0 (3):\penalty0 556--559, 1985.

\bibitem[Rikitake(1980)]{Rikitake_80}
H.~Tanaka N. Ohshiman Y. Sasai Y. Ishikawa S. Koyama M. Kawamura e K.~Ohchi
  T.~Rikitake, Y.~Honkura.
\newblock Changes in the geomagnetic field associated with earthquakes in the
  izu peninsula, japan.
\newblock \emph{J. Geomag. Geoelectr.}, 32:\penalty0 721--739, 1980.

\end{thebibliography}

\end{document}